---
title: "P8105_HW1_zw3268"
author: "Zeheng Wei"
date: "2026-09-24"
output: github_document
---
```{r}
library(tidyverse)
```
## Problem_1 

Importing the dataset 
```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset includes information about penguins’ species, island, sex, and year, along with measurements of bill length, bill depth, flipper length, and body mass. The dataset includes three species (Adelie, Chinstrap, and Gentoo) and three islands (Biscoe, Dream, and Torgersen). It contains `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE) %>% round(3)` mm.

```{r}
penguins %>% drop_na() %>% 
ggplot(aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
geom_point()
```

Export scatter plot
```{r}
ggsave("penguin_plot.jpg")
```

## Problem_2

Creating the data frame: 

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”
```{r}
df = tibble(random_sample = rnorm(10), 
            logical_vector = random_sample > 0, 
            character = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
            factor = factor(c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A"))
)
df
```

Taking the mean of each vector
```{r}
mean(pull(df, random_sample))
mean(pull(df, logical_vector))
mean(pull(df, character))
mean(pull(df, factor))
```
The mean can be calculated for the numeric and logical vectors. For the logical vector, TRUE is treated as 1 and FALSE as 0, so its mean represents the proportion of sampled values greater than zero. The character and factor vectors return NA with warnings because they are neither numeric nor logical.


Applying as.numeric() to the logical vector converts TRUE to 1 and FALSE to 0, explaining why its mean equals the proportion of sampled values greater than zero.
```{r, results='hide', warning=FALSE}
as.numeric(pull(df, logical_vector))
as.numeric(pull(df, character))
as.numeric(pull(df, factor))
```
The character vector contains letters that cannot be interpreted as numbers, so conversion produces NA values and a warning. The factor vector is converted to its internal level codes: 1 for A, 2 for B, and 3 for C. These codes represent categories rather than meaningful quantities. Although factors have underlying numeric codes, mean() does not directly accept factor or character vectors, so their means return NA with warnings.
